---
title: "Untitled"
author: "JJayes"
date: "17/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(readxl)
```

## Purpose

Group project for EKHM63 - Growth over space and time

### Reading in data

```{r}
df <- read_excel(here("data", "USA.xlsx")) %>% 
  janitor::clean_names()
```

### Tidy to long format
```{r}
df <- df %>% 
  pivot_longer(-c(indicator, measure), names_to = "year") %>% 
  mutate(year = parse_number(year))
```

### Create labour productivity

```{r}
df_labour_prod <- df %>% 
  filter(indicator %in% c("GDP (volume)", "Total Hours")) %>% 
  select(indicator, year, value) %>% 
  pivot_wider(names_from = indicator) %>% 
  mutate(labour_productivity = `GDP (volume)` / `Total Hours`) %>% 
  select(year, labour_productivity)

df_labour_prod <- df_labour_prod %>% 
  mutate(indicator = "Labour Productivity",
         measure = "`GDP (volume)` / `Total Hours`") %>% 
  rename(value = labour_productivity)

df <- df %>% 
  bind_rows(df_labour_prod)
```



```{r}
df %>% 
  distinct(indicator)
```

### Calculate growth rates

```{r}
df <- df %>% 
  group_by(indicator) %>% 
  mutate(growth_rate = (value - lag(value))/lag(value)*100) %>% 
  ungroup()
```

### Write out data

```{r}
# df %>% 
#   mutate(country = "USA") %>% 
#   write_rds(here("data", "df_usa.rds"))
```

### Read in new data

```{r}
# df_long <- df_brazil %>% 
#   bind_rows(df_france) %>% 
#   bind_rows(df_germany) %>% 
#   bind_rows(df_india) %>% 
#   bind_rows(df_italy) %>% 
#   bind_rows(df_japan) %>% 
#   bind_rows(df_singapore) %>% 
#   bind_rows(df_spain) %>% 
#   bind_rows(df_sweden) %>% 
#   bind_rows(df_uk) %>% 
#   bind_rows(df_usa) 

# df_long %>% write_rds(here("df_all_countries.rds"), compress = "gz")
```

### Read in all

```{r}
df <- read_rds(here("data", "df_all_countries.rds"))
```


```{r}
df %>% 
  filter(str_detect(indicator, "Employment")) %>% 
  pivot_longer(cols = c(value, growth_rate)) %>% 
  ggplot(aes(year, value)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = scales::number_format()) +
  facet_wrap(~ name, scales = "free_y", nrow = 2)
```

```{r}
tc <- 5

df_time_chunks <- df %>% 
  group_by(indicator) %>% 
  mutate(time_chunk = year - year %% tc) %>% 
  ungroup() %>% 
  group_by(indicator, time_chunk) %>% 
  mutate(index = row_number()) %>% 
  filter(index %in% c(1, tc)) %>% 
  mutate(ratio_value = value / lag(value),
         one_over_change_year = 1/(year - lag(year)),
         compound_gr = ratio_value^(one_over_change_year)-1) %>% 
  ungroup() 

df_time_chunks %>% 
  filter(str_detect(indicator, "Total Hours"),
         !is.na(compound_gr)) %>% 
  ggplot(aes(year, compound_gr)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = scales::percent_format())
```


GDP (volume), Total Hours, Capital Stock

```{r}
df_time_chunks %>% 
  filter(indicator %in% c("GDP (volume)", "Total Hours", "Capital Stock (K)"),
         !is.na(compound_gr)) %>% 
  ggplot(aes(year, compound_gr, colour = indicator)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, lty = 2) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(y = "Compound Growth Rate",
       x = NULL,
       colour = "Indicator")
```


```{r}
# capital weight
alpha <- .3
# labour weight
beta <- .7
```


```{r}

df_contributions <- df_time_chunks %>% 
  filter(indicator %in% c("GDP (volume)", "Total Hours", "Capital Stock (K)"),
         !is.na(compound_gr)) %>% 
  select(indicator, year, compound_gr) %>% 
  pivot_wider(names_from = indicator, values_from = compound_gr) %>% 
  janitor::clean_names() %>% 
  mutate(capital_contribution = alpha*capital_stock_k,
         labour_contribution = beta*total_hours,
         crude_tfp = gdp_volume - capital_contribution - labour_contribution) %>% 
  pivot_longer(-year) %>% 
  filter(name %in% c("capital_contribution", "labour_contribution", "crude_tfp", "gdp_volume"))

df_contributions %>% 
  pivot_wider(names_from = name) %>% 
  pivot_longer(-c(year, gdp_volume)) %>% 
  ggplot() +
  geom_col(aes(year, value, fill = name), 
           position = "dodge", 
           alpha = .8) +
  geom_point(aes(year, gdp_volume, colour = "GDP Growth")) +
  geom_line(aes(year, gdp_volume, colour = "GDP Growth")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = NULL,
       fill = "Contribution to GDP Growth",
       colour = "Level of GDP Growth")
```



```{r}
df_time_chunks %>% 
  filter(indicator %in% c("GDP (volume)", "Total Hours", "Capital Stock (K)", "Labour Productivity"),
         !is.na(compound_gr)) %>% 
  select(indicator, year, compound_gr) %>% 
  pivot_wider(names_from = indicator, values_from = compound_gr) %>% 
  janitor::clean_names() %>% 
  mutate(k_per_worker_growth = capital_stock_k - total_hours,
         contribution_of_capital_deepening = k_per_worker_growth*alpha,
         crude_tfp = labour_productivity - contribution_of_capital_deepening) %>% 
  pivot_longer(-year) %>% 
  filter(name %in% c("gdp_volume", "labour_productivity",  "crude_tfp", "contribution_of_capital_deepening")) %>% 
  pivot_wider(names_from = name) %>% 
  pivot_longer(-c(year, gdp_volume)) %>% 
  mutate(name = str_to_sentence(str_replace_all(name, "_", " "))) %>% 
  ggplot() +
  geom_col(aes(year, value, fill = name), position = "dodge") +
  geom_point(aes(year, gdp_volume, colour = "GDP Growth")) +
  geom_line(aes(year, gdp_volume, colour = "GDP Growth")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = NULL,
       y = "GDP growth rate\n contribution to growth",
       colour = "Level of GDP Growth",
       fill = "Contribution to GDP Growth")

```

### Human capital contribution.

```{r}
df_calcs <- df_time_chunks %>%
  filter(
    indicator %in% c("Human Capital Stock (HK)", "GDP (volume)", "Total Hours", "Capital Stock (K)", "Labour Productivity"),
    !is.na(compound_gr)
  ) %>%
  select(indicator, year, compound_gr) %>%
  pivot_wider(names_from = indicator, values_from = compound_gr) %>%
  janitor::clean_names() %>%
  mutate(
    # capital
    capital_contribution = alpha * capital_stock_k,
    # labour
    labour_contribution = beta * total_hours,
    # human capital
    contribution_of_human_capital = human_capital_stock_hk * beta,
    # capital per worker growth
    k_per_worker_growth = capital_stock_k - total_hours,
    # capital deepening
    contribution_of_capital_deepening = k_per_worker_growth * alpha,
    # crude tfp via labour prod
    crude_tfp_via_labour_productivity = labour_productivity - contribution_of_capital_deepening,
    # crude tfp via gdp
    crude_tfp_via_gdp = gdp_volume - capital_contribution - labour_contribution,
    # now here we are just doing the method based on GDP growth
    labour_augmented_tfp_via_gdp = crude_tfp_via_gdp - contribution_of_human_capital,
    # here we do method based on labour productiviy
    labour_augmented_tfp_via_labour_productivity = crude_tfp_via_labour_productivity - contribution_of_human_capital
  ) %>% 
  pivot_longer(-year)

```

Must make a plot of ICT and non-ICT stock and contribution to growth.

- Point is to show how even though ICT stock is small, it's importance to growth is large

### ICT

Plot of ICT revolution

```{r}
df_time_chunks %>%
  filter(
    str_detect(indicator, "ICT Capital"),
    !is.na(compound_gr)
  ) %>%
  mutate(share = if_else(str_detect(indicator, "Share"), "Share of capital stock", "Growth in capital stock"),
         ict = if_else(str_detect(indicator, "Non"), "Non ICT", "ICT")) %>%
  ggplot(aes(time_chunk, value, colour = ict, group = indicator)) +
  geom_point() +
  geom_line() +
  facet_wrap(~share, scales = "free_y", nrow = 2) +
  labs(x = NULL,
       y = NULL,
       colour = "Sector")

```


```{r}

df_ict <- df %>% 
  group_by(indicator) %>% 
  mutate(time_chunk = year - year %% tc) %>% 
  ungroup() %>% 
  group_by(indicator, time_chunk) %>% 
  mutate(mean_in_tc = mean(value, na.rm = T)) %>% 
  ungroup() %>% 
  filter(str_detect(indicator, "ICT Capital"),
         !is.na(mean_in_tc)) %>% 
  distinct(time_chunk, mean_in_tc, .keep_all = T)

df_ict %>% 
  ggplot(aes(year, mean_in_tc, colour = indicator)) +
  geom_point() +
  geom_line()

df_time_chunks %>% 
  filter(!is.na(compound_gr),
         str_detect(indicator, "ICT Capital")) %>% 
  select(indicator, time_chunk, compound_gr)

```

We need to multiply ICT capital growth by the average ICT capital share in the time chunk to get the ICT capital share contribution.

From df_ict we get 
```{r}
df_ict_joined <- df %>% 
  group_by(indicator) %>% 
  mutate(time_chunk = year - year %% tc) %>% 
  ungroup() %>% 
  group_by(indicator, time_chunk) %>% 
  mutate(mean_in_tc = mean(value, na.rm = T)) %>% 
  ungroup() %>% 
  filter(str_detect(indicator, "ICT Capital"),
         !is.na(mean_in_tc)) %>% 
  distinct(time_chunk, mean_in_tc, .keep_all = T) %>% 
  select(indicator, time_chunk, mean_in_tc) %>% 
  inner_join(df_time_chunks %>% 
  filter(!is.na(compound_gr),
         str_detect(indicator, "ICT Capital")) %>% 
  select(indicator, time_chunk, compound_gr),
  by = c("indicator", "time_chunk")) %>% 
  mutate(compound_gr = case_when(
    
    str_detect(indicator, "Volume") ~ compound_gr * 100,
    TRUE ~ compound_gr
    
  )) %>% 
   mutate(mean_in_tc = case_when(
    
    str_detect(indicator, "Share") ~ mean_in_tc / 100,
    TRUE ~ mean_in_tc
    
  )) %>% 
  mutate(value = case_when(
    
    str_detect(indicator, "Share") ~ mean_in_tc,
    TRUE ~ compound_gr
    
  )) %>% 
  select(indicator, time_chunk, value) %>% 
  pivot_wider(names_from = indicator) %>% 
  janitor::clean_names()

df_ict_joined %>% 
  mutate(non_ict_capital_contribution = non_ict_capital_volume * non_ict_capital_share,
         ict_capital_contribution = ict_capital_volume * ict_capital_share) %>% 
  pivot_longer(-time_chunk) %>% 
  mutate(sector = ifelse(str_detect(name, "non"), "Non-ICT", "ICT"),
         type = case_when(
           
           str_detect(name, "contribution") ~ "Contribution to GDP growth",
           str_detect(name, "share") ~ "Share of capital",
           TRUE ~ "Growth of capital"
           
         )) %>% 
  ggplot(aes(time_chunk, value, colour = type)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ sector) +
  labs(x = NULL)

```


### Range calculations

```{r}

year_1 <- 1950
year_2 <- 1960

df_range_1 <- df %>%
  filter(
    country == "France",
    year %in% c(year_1, year_2)
  ) %>%
  group_by(indicator) %>%
  mutate(
    ratio_value = value / lag(value),
    one_over_change_year = 1 / (year - lag(year)),
    compound_gr = (ratio_value^(one_over_change_year) - 1)*100,
    range = paste0(year_1, "-", year_2)
  ) %>%
  ungroup() %>% 
  select(indicator, year, value, country, compound_gr, range) %>% 
  filter(!is.na(compound_gr)) %>% 
  select(indicator, range, compound_gr) %>% 
  pivot_wider(names_from = indicator, values_from = compound_gr) %>% 
  janitor::clean_names() %>% 
  mutate( 
    capital_contribution = alpha * capital_stock_k,
    labour_contribution = beta * total_hours,
    contribution_of_human_capital = human_capital_stock_hk * beta,
    k_per_worker_growth = capital_stock_k - total_hours,
    contribution_of_capital_deepening = k_per_worker_growth * alpha,
    crude_tfp_via_labour_productivity = labour_productivity - contribution_of_capital_deepening,
    crude_tfp_via_gdp = gdp_volume - capital_contribution - labour_contribution,
    labour_augmented_tfp_via_gdp = crude_tfp_via_gdp - contribution_of_human_capital,
    labour_augmented_tfp_via_labour_productivity = crude_tfp_via_labour_productivity - contribution_of_human_capital) %>% 
  pivot_longer(-range)

df_range_1 %>% 
  filter(name %in% c("capital_contribution", "labour_contribution", "crude_tfp_via_gdp", "gdp_volume")) %>% 
  pivot_wider(names_from = name) %>% 
  pivot_longer(-c(range, gdp_volume)) %>% 
  ggplot() +
  geom_col(aes(range, value, fill = name), alpha = .5) +
  geom_point(aes(range, gdp_volume, colour = "GDP Growth"))
  
  

```

